---
title: RNA-Seq Analysis – 
 Differential Gene Expression and Consistent Gene Differential Expression analysis
author: "Beatriz Manso"
date: '2022-04-21'
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
    code_folding: false
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Find differentially expressed genes in PLancenta tissue with preeclampsia 
We will need the control samples for normal placenta tissue and samples of tissue with preeclampsia . 

# Methods

Set working directory:

```{r}
setwd("C:/Users/manso/OneDrive - University of West London/MSc Bioinformatics - UWL/6.BGA - Bioinformatics and Genome Analysis/week 5 - Microarray analysis/practical")
```

Install necessary packages and load libraries:

```{r}
# Installing DEseq2 package
if (!require("BiocManager", quietly = TRUE))
 install.packages("BiocManager")
BiocManager::install("DESeq2")

# Load the Library
library(DESeq2)
library(tidyverse)
```
# Part 1: Differential Gene Expression Analysis

## 1. Read Data

```{r}

countsN <- read.csv("ExpData.csv")

rownames(countsN) <- countsN[,1]

countsN <- countsN[,-1]

dim(countsN)
```
In this dataset we have 16 samples ans 60715 gene ids.

```{r}
#Condition and Coldata 
condition <- as.factor(c(rep("NP", 8,), rep("PE", 8)))

colData <- data.frame(row.names=colnames(countsN), 
                      condition=factor(condition, levels=c('NP', 'PE')))

```

## 2. DEseqDataset

```{r}
#DEseqDataset
dataset <- DESeqDataSetFromMatrix(countData = countsN,
                                  colData = colData, 
                                  design = ~condition)
dataset
```
## 3. Run the DESeq2 algorithm and extract results for our two-class

```{r}
dds <- DESeq(dataset)
```

```{r}
print(dds)
```
## 4. Get Results

```{r}
result <- results(dds, contrast=c('condition','NP','PE'), alpha = 0.01)
result <- results(dds, contrast=c('condition','NP','PE'), alpha = 0.01)
summary(result)

```
## 5. Extract the up and down regulated gene file

```{r}
resSigind = result[ which(result$padj < 0.01 & result$log2FoldChange > 0), ]
resSigrep = result[ which(result$padj < 0.01 & result$log2FoldChange < 0), ]
resSig = rbind(resSigind, resSigrep)

write.csv(resSig, "Week9_DGEUp&DnReg0.01.csv")
```


# Part 2: Consistent Gene Differential Expression (CGDE)

```{r}
if (!require("BiocManager", quietly = TRUE))
 install.packages("BiocManager")
BiocManager::install("Rmpfr")
BiocManager::install("RankProd")
BiocManager::install("Rmpfr")


library(edgeR)
library(RankProd)

```

## 1. Read Data

```{r}
countsN <- read.csv("ExpData.csv", header = T)
rownames(countsN) <- countsN[,1]
countsN <- countsN[,-1]
countsN <- as.matrix(countsN)
dim(countsN)
```
## 2. RNA-Seq Normalisation with TMM from edgeR Package
```{r}
dge <- DGEList(counts=countsN)
dge <- calcNormFactors(dge, method = "TMM")
TPMnorm_NPET1 <- cpm(dge, log = TRUE, normalized.lib.sizes=TRUE)

dim(countsN)

```

## 3. Class

```{r}
n1 <- 8
n2 <- 8
cl <- c(rep(0,n1), rep(1,n2))
```


## 4. Origin

```{r}
origin <- rep(1, 13) # If origin is single, origin information is not needed for RP analysis. 
```

## 5. RP

```{r}
RP <- RP(countsN, cl, num.perm=100,
         logged=TRUE, na.rm=FALSE, 
         gene.names=rownames (countsN),
         plot=F, rand=123)

```

## 6. Top genes

```{r}
TopGene <- topGene(RP, cutoff=0.001, method="pfp",
                   logged=TRUE, logbase=2, gene.names=rownames(countsN))

```


```{r}
T1 <- TopGene[["Table1"]]
T2 <- TopGene[["Table2"]]

```

7. Extract the up and down regulated genes file

```{r}
write.table(T1 , col.names=NA, row.names=T, file ="Week9_CGDEUpReg0.01.tsv", sep ="\t")
write.table(T2 , col.names=NA, row.names=T, file ="Week9_CGDEDnReg0.01.tsv", sep ="\t")

dim(T1)
```

```{r}
dim(T2)
```










